An equivalence of compositional frameworks

### 2019-01-04 EquivCompFrameworks_20180911kc.tex
Polishing Theorem 2.2

Let $\mathrm{A}$ be a category with finite colimits and $F \colon \mathrm{A} \to \mathbf{Cat}$ a symmetric lax monoidal pseudofunctor. Then the double category $\mathbb{F}\textnormal{Cospan}(\mathrm{A})$ is symmetric monoidal.

@@

First we note that the category of objects $\mathbb{F}\textnormal{Cospan}(\mathrm{A})_0=\mathrm{A}$ is symmetric monoidal under binary coproducts and the left and right unitors, associators and braidings are given as natural maps. The category of arrows $\mathbb{F}\textnormal{Cospan}(\mathrm{A})_1$ has:

@@

\item{objects as pairs:
\node (D) at (3,0) {$d \in F(c)$};

@@

\item{morphisms as maps of cospans in $\mathrm{A}$

@@

together with elements of the images of the apices under the pseudofunctor $F$ and a morphism $\iota \colon F(h)(d) \to d^\prime$.

@@

Given two objects $M_1$ and $M_2$ of $\mathbb{F}\textnormal{Cospan}(\mathrm{A})_1$:

@@

their tensor product $M_1 \otimes M_2$ is given by taking the coproducts of the cospans of $\mathrm{A}$
\node (B) at (1.25,0) {$c_1+c_2$};
\node (C) at (2.5,0) {$b_1+b_2$};
\node (D) at (4.25,0) {$d_{M_1 \otimes M_2} \in F(c_1+c_2)$};
and where the element of the image of the apex under $F$  is obtained using the natural transformation of the symmetric lax monoidal pseudofunctor $F:$ $$1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{d_1 \times d_2} F(c_1) \times F(c_2) \xrightarrow{\phi_{c_1,c_2}} F(c_1 + c_2).$$The monoidal unit is given by:

@@

\node (D) at (3,0) {$! \in F(0)$};

@@

where $0$ is the monoidal unit of $\mathrm{A}$ and $! \colon 1 \to F(0)$ is the morphism which is part of the structure of the symmetric lax monoidal pseudofunctor $F \colon \mathrm{A} \to \mathbf{Cat}$. Tensoring an object with the monoidal unit, say, on the left:

@@

\node (D) at (1,-0.5) {$! \in F(0)$};
\node (E) at (3,0) {$a$};
\node (F) at (4,0) {$c$};
\node (H) at (4,-0.5) {$d \in F(c)$};

@@

\node (A) at (0,0) {$0+a$};
\node (B) at (1,0) {$0+c$};
\node (C) at (2,0) {$0+b$};
\node (D) at (3.5,0) {$d^\prime \in F(0+c)$};

@@

where $d^\prime \in F(0+c)$ is given by $$1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{\phi \times d} F(0) \times F(c) \xrightarrow{\phi_{0,d}} F(0+c).$$The left unitor is then a morphism in $\mathbb{F} \textnormal{Cospan}(\mathrm{A})_1$ given by:
\node (A) at (0,0.5) {$0+a$};
\node (A') at (0,-0.5) {$a$};
\node (B) at (1,0.5) {$0+c$};
\node (C) at (2,0.5) {$0+b$};
\node (C') at (2,-0.5) {$b$};
\node (D) at (1,-0.5) {$c$};
\node (E) at (3.5,0.5) {$d^\prime \in F(0+c)$};
\node (F) at (3.5,-0.5) {$d \in F(c)$};

@@

where $\ell$ is the left unitor of $(\mathrm{A},+,0)$, together with the morphism $\iota_{\lambda} \colon F(\ell)(d^\prime) \to d$. The right unitor is similar.

Given three objects $M_1, M_2$ and $M_3$ in $\mathbb{F}\textnormal{Cospan}(\mathrm{A})_1$:

@@

@@

\node (L) at (7,-0.5) {$d_3 \in F(c_3)$};

@@

where $d_{(M_1 \otimes M_2) \otimes M_3} \colon 1 \to F((c_1+c_2)+c_3)$ is given by: $$1 \xrightarrow{((d_1 \times d_2) \times d_3)} (F(c_1) \times F(c_2)) \times F(c_3) \xrightarrow{\phi_{c_1,c_2} \times 1} F(c_1+c_2) \times F(c_3) \xrightarrow{\phi_{c_1+c_2,c_3}} F((c_1+c_2)+c_3)$$whereas tensoring the last two and then the first results in $M_1 \otimes (M_2 \otimes M_3)$:

@@

where $d_{M_1 \otimes (M_2 \otimes M_3)} \colon 1 \to F(c_1+(c_2+c_3))$ is given by: $$1 \xrightarrow{d_1 \times (d_2 \times d_3)} F(c_1) \times (F(c_2) \times F(c_3)) \xrightarrow{1 \times \phi_{c_2,c_3}} F(c_1) \times F(c_2+c_3) \xrightarrow{\phi_{c_1,c_2+c_3}} F(c_1+(c_2+c_3)).$$If we let $a$ denote the associator of $(\mathrm{A},+,0)$, the associator of $\mathbb{F} \textnormal{Cospan}(\mathrm{A})_1$ is then a map of cospans in $\mathrm{A}$ from $(M_1 \otimes M_2) \otimes M_3)$ to $M_1 \otimes (M_2 \otimes M_3)$ given by:

@@

together with the isomorphism $\iota_a \colon F(a)(d_{(M_1 \otimes M_2) \otimes M_3}) \to d_{M_1 \otimes (M_2 \otimes M_3)}$. If we denote the above associator simply as $a$ and the left and right unitors as $\lambda$ and $\rho$, respectively, then given four objects in $\mathbb{F}\textnormal{Cospan}(\mathrm{A})_1$, say $M_1, M_2, M_3$ and $M_4$:

@@

\node (L) at (1,-2) {$d_3 \in F(c_3)$};

@@

\node (P) at (4,-2) {$d_4 \in F(c_4)$};

@@

then the following pentagon of underlying cospans commutes:

@@

since $$F(aa)(d_{((M_1 \otimes M_2) \otimes M_3) \otimes M_4})=F((1 \otimes a)a(a \otimes 1))(d_{((M_1 \otimes M_2) \otimes M_3) \otimes M_4})$$ as the above underlying diagram of maps of cospans commutes.

Similarly, if we denote the left and right unitors as $\lambda$ and $\rho$, respectively, then the following triangle of underlying maps of cospans commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(M_1 \otimes U_{0}) \otimes M_2$};
\node (B) at (2.25,1) {$M_1 \otimes M_2$};
\node (C) at (4.5,0) {$M_1 \otimes (U_0 \otimes M_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[left]{$\rho \otimes 1$} (B)
(C) edge node[right]{$1 \otimes \lambda$} (B)
(A) edge node[above]{$a$} (C);
\end{tikzpicture}
\]
as well as the following triangle of corresponding decorations in the category $F(c_1+c_2)$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$F(\rho \otimes 1)(d_{(M_1 \otimes U_0) \otimes M_2})$};
\node (B) at (2.25,1) {$d_{M_1 \otimes M_2}$};
\node (C) at (4.5,0) {$F(1 \otimes \lambda)(d_{M_1 \otimes (U_0 \otimes M_2)}) $};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[left]{$\iota_{\rho \otimes 1}$} (B)
(C) edge node[right]{$\iota_{1 \otimes \lambda}$} (B)
(A) edge node[above]{$F(\iota_a)$} (C);
\end{tikzpicture}
\]
since $$F(\rho \otimes 1)(d_{(M_1 \otimes U_0) \otimes M_2})=F(a \otimes (1 \otimes \lambda))(d_{(M_1 \otimes U_0) \otimes M_2})$$ as the above underlying diagram of maps of cospans commutes.

For a monoidal product of objects $M_1 \otimes M_2$ in $\mathbb{F} \textnormal{Cospan}(\mathrm{A})_1$, the source and target structure functors $S,T \colon \mathbb{F} \textnormal{Cospan}(\mathrm{A})_1 \to \mathbb{F} \textnormal{Cospan}(\mathrm{A})_0$ satisfy the following equations: $$S(M_1 \otimes M_2)=S(M_1) \otimes S(M_2)$$ $$T(M_1 \otimes M_2)=T(M_1) \otimes T(M_2).$$
For two objects $M_1$ and $M_2$ in $\mathbb{F} \textnormal{Cospan}(\mathrm{A})_1$, we have a braiding $\beta_{M_1,M_2} \colon M_1 \otimes M_2 \to M_2 \otimes M_1$ given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a_1+a_2$};
\node (A') at (0,-0.5) {$a_2+a_1$};
\node (B) at (2,0.5) {$c_1+c_2$};
\node (C) at (4,0.5) {$b_1+b_2$};
\node (C') at (4,-0.5) {$b_2+b_1$};
\node (D) at (2,-0.5) {$c_2+c_1$};
\node (E) at (5.5,0.5) {$d_{M_1 \otimes M_2} \in F(c_1+c_2)$};
\node (F) at (5.5,-0.5) {$d_{M_2 \otimes M_1} \in F(c_2+c_1)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1+i_2$} (B)
(C) edge node[above]{$o_1+o_2$} (B)
(A) edge node[left]{$\beta_{a_1,a_2}$} (A')
(C) edge node[left]{$\beta_{b_1,b_2}$} (C')
(A') edge node [above]{$i_2+i_1$} (D)
(C') edge node [above]{$o_2+o_1$} (D)
(B) edge node [left] {$\beta_{c_1,c_2}$} (D);
\end{tikzpicture}
\]
$$\iota_{\beta_{M_1,M_2}} \colon F(\beta_{c_1,c_2})(d_{M_1 \otimes M_2}) \xrightarrow{\sim} d_{M_2 \otimes M_1}.$$
This braiding makes the following triangle of underlying cospans commute:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$M_1 \otimes M_2$};
\node (B) at (2.25,1) {$M_1 \otimes M_2$};
\node (C) at (4.5,0) {$M_2 \otimes M_1$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$1$} (B)
(C) edge node[above]{$\beta_{M_2,M_1}$} (B)
(A) edge node[above]{$\beta_{M_1,M_2}$} (C);
\end{tikzpicture}
\]
as well as the following diagram of corresponding decorations in the category $F(c_2+c_1)$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$F(\beta_{c_1,c_2})(d_{M_1 \otimes M_2})$};
\node (B) at (2.25,1) {$F(\beta_{c_1,c_2})(d_{M_1 \otimes M_2}) $};
\node (C) at (4.5,0) {$d_{M_2 \otimes M_1} $};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$1$} (B)
(C) edge node[right]{$F(\beta_{c_1,c_2})(\iota_{\beta_{M_2,M_1}})$} (B)
(A) edge node[above]{$\iota_{\beta_{M_1,M_2}}$} (C);
\end{tikzpicture}
\]
since $F(\beta_{c_1,c_2} \beta_{c_2,c_1}) (d_{M_2 \otimes M_1}) = d_{M_2 \otimes M_1}$. Thus $\mathbb{F} \textnormal{Cospan}(\mathrm{A})_1$ is also symmetric monoidal.

>>>>

Let $\mathsf{A}$ be a category with finite colimits and $F \colon \mathsf{A} \to \mathbf{Cat}$ a symmetric lax monoidal pseudofunctor. Then the double category $F\mathbb{C}$sp is symmetric monoidal.

@@

First we note that the category of objects $F\mathbb{C}\textnormal{sp}_0=\mathsf{A}$ is symmetric monoidal under binary coproducts and the left and right unitors, associators and braidings are given as natural maps. The category of arrows $F\mathbb{C}\textnormal{sp}_1$ has:

@@

\item{objects as $F$-decorated cospans which are pairs:

@@

\item{morphisms as maps of cospans in $\mathsf{A}$

@@

together with a morphism $\iota \colon F(h)(d) \to d^\prime$.

@@

Given two objects $M_1$ and $M_2$ of $F\mathbb{C}\textnormal{sp}_1$:

@@

\node (E) at (3,0) {$a_1^\prime$};
\node (G) at (5,0) {$a_2^\prime$};
(G) edge node[above]{$o^\prime$} (F);
their tensor product $M_1 \otimes M_2$ is given by taking the coproducts of the cospans of $\mathsf{A}$
\node (A) at (0,0) {$a_1+a_1^\prime$};
\node (B) at (1.25,0) {$b+b^\prime$};
\node (C) at (2.5,0) {$a_2+a_2^\prime$};
\node (D) at (4.25,0) {$d_{M_1 \otimes M_2} \in F(b+b')$};
(A) edge node[above]{$i+i'$} (B)
(C) edge node[above]{$o+o'$} (B);
and where the decoration on the apex is obtained using the natural transformation of the symmetric lax monoidal pseudofunctor $F:$ $$d_{M_1 \otimes M_2} \coloneqq 1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{d_{M_1} \times d_{M_2}} F(b) \times F(b') \xrightarrow{\phi_{b,b'}} F(b+b').$$The monoidal unit 0 is given by:

@@

\node (D) at (3,0) {$!_0 \in F(0)$};

@@

where $0$ is the monoidal unit of $\mathsf{A}$ and $!_0 \colon 1 \to F(0)$ is the morphism which is part of the structure of the symmetric lax monoidal pseudofunctor $F \colon \mathsf{A} \to \mathbf{Cat}$. Tensoring an object with the monoidal unit, say, on the left:

@@

\node (D) at (1,-0.5) {$!_0 \in F(0)$};
\node (E) at (3,0) {$a_1$};
\node (H) at (4,-0.5) {$d_M \in F(b)$};
\node (I) at (2.5,0) {$\otimes$};

@@

\node (A) at (0,0) {$0+a_1$};
\node (B) at (1,0) {$0+b$};
\node (C) at (2,0) {$0+a_2$};
\node (D) at (3.5,0) {$d_{0 \otimes M} \in F(0+b)$};

@@

where $d_{0\otimes M} \in F(0+b)$ is given by $$1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{!_0 \times d_M} F(0) \times F(b) \xrightarrow{\phi_{0,b}} F(0+b).$$The left unitor is then an isomorphism in $F\mathbb{C}\textnormal{sp}_1$ given by:
\node (A) at (0,0.5) {$0+a_1$};
\node (B) at (1,0.5) {$0+b$};
\node (C) at (2,0.5) {$0+a_2$};
\node (C') at (2,-0.5) {$a_2$};
\node (D) at (1,-0.5) {$b$};
\node (E) at (3.5,0.5) {$d_{0+M} \in F(0+b)$};
\node (F) at (3.5,-0.5) {$d_M \in F(b)$};

@@

where $\ell$ is the left unitor of $(\mathsf{A},+,0)$, together with the isomorphism $\iota_{\lambda} \colon F(\ell)(d_{0 \otimes M}) \to d_M$ which is specified by the isomorphism of categories $F(0+b) \cong F(b)$. The right unitor is similar.

Given three objects $M_1, M_2$ and $M_3$ in $F\mathbb{C}\textnormal{sp}_1$:

@@

\node (H) at (4,-0.5) {$d_{M_2} \in F(c_2)$};

@@

\node (L) at (7,-0.5) {$d_{M_3} \in F(c_3)$};

@@

where $d_{(M_1 \otimes M_2) \otimes M_3} \colon 1 \to F((c_1+c_2)+c_3)$ is given by: $$1 \xrightarrow{(d_{M_1} \times d_{M_2}) \times d_{M_3})} (F(c_1) \times F(c_2)) \times F(c_3) \xrightarrow{\phi_{c_1,c_2} \times 1} F(c_1+c_2) \times F(c_3) \xrightarrow{\phi_{c_1+c_2,c_3}} F((c_1+c_2)+c_3)$$whereas tensoring the last two and then the first results in $M_1 \otimes (M_2 \otimes M_3)$:

@@

where $d_{M_1 \otimes (M_2 \otimes M_3)} \colon 1 \to F(c_1+(c_2+c_3))$ is given by: $$1 \xrightarrow{d_{M_1} \times (d_{M_2} \times d_{M_3})} F(c_1) \times (F(c_2) \times F(c_3)) \xrightarrow{1 \times \phi_{c_2,c_3}} F(c_1) \times F(c_2+c_3) \xrightarrow{\phi_{c_1,c_2+c_3}} F(c_1+(c_2+c_3)).$$If we let $a$ denote the associator of $(\mathsf{A},+,0)$, the associator of $F\mathbb{C}\textnormal{sp}_1$ is then a map of cospans in $\mathsf{A}$ from $(M_1 \otimes M_2) \otimes M_3$ to $M_1 \otimes (M_2 \otimes M_3)$ given by:

@@

together with the isomorphism $\iota_a \colon F(a)(d_{(M_1 \otimes M_2) \otimes M_3}) \to d_{M_1 \otimes (M_2 \otimes M_3)}$. If we denote the above associator simply as $a$ and the left and right unitors as $\lambda$ and $\rho$, respectively, then given four objects in $F\mathbb{C}\textnormal{sp}_1$, say $M_1, M_2, M_3$ and $M_4$:

@@

\node (L) at (1,-2) {$d_{M_3} \in F(c_3)$};

@@

\node (P) at (4,-2) {$d_{M_4} \in F(c_4)$};

@@

then as $\mathbb{C}\textnormal{sp}(\mathsf{A})$ is a symmetric monoidal double category, the following pentagon of underlying cospans commutes:

@@

since $$F(aa)(d_{((M_1 \otimes M_2) \otimes M_3) \otimes M_4})=F((1 \otimes a)a(a \otimes 1))(d_{((M_1 \otimes M_2) \otimes M_3) \otimes M_4})$$ as the corresponding pentagon in the symmetric monoidal category $(\mathsf{A},+,0)$ commutes.

Similarly, if we denote the left and right unitors as $\lambda$ and $\rho$, respectively, then the following triangle of underlying maps of cospans commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(M_1 \otimes 0) \otimes M_2$};
\node (B) at (2.25,1) {$M_1 \otimes M_2$};
\node (C) at (4.5,0) {$M_1 \otimes (0 \otimes M_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$\rho \otimes 1$} (B)
(C) edge node[above]{$1 \otimes \lambda$} (B)
(A) edge node[above]{$a$} (C);
\end{tikzpicture}
\]
as well as the following triangle of corresponding decorations in the category $F(c_1+c_2)$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$F(\rho \otimes 1)(d_{(M_1 \otimes 0) \otimes M_2})$};
\node (B) at (2.25,1) {$d_{M_1 \otimes M_2}$};
\node (C) at (4.5,0) {$F(1 \otimes \lambda)(d_{M_1 \otimes (0 \otimes M_2)}) $};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[left]{$\iota_{\rho \otimes 1}$} (B)
(C) edge node[right]{$\iota_{1 \otimes \lambda}$} (B)
(A) edge node[above]{$F(1 \otimes \lambda)(\iota_a)$} (C);
\end{tikzpicture}
\]
since $$F(\rho \otimes 1)(d_{(M_1 \otimes 0) \otimes M_2})=F((1 \otimes \lambda)a)(d_{(M_1 \otimes 0) \otimes M_2})$$ as the corresponding triangle in the symmetric monoidal category $(\mathsf{A},+,0)$ commutes.

For a tensor product of objects $M_1 \otimes M_2$ in $F\mathbb{C}\textnormal{sp}_1$, the source and target structure functors $S,T \colon F\mathbb{C}\textnormal{sp}_1 \to F\mathbb{C}\textnormal{sp}_0$ satisfy the following equations: $$S(M_1 \otimes M_2)=S(M_1) \otimes S(M_2)$$ $$T(M_1 \otimes M_2)=T(M_1) \otimes T(M_2).$$
For two objects $M_1$ and $M_2$ in $F\mathbb{C}\textnormal{sp}_1$, we have a braiding $\beta_{M_1,M_2} \colon M_1 \otimes M_2 \to M_2 \otimes M_1$ given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a_1+a_2$};
\node (A') at (0,-0.5) {$a_2+a_1$};
\node (B) at (2,0.5) {$c_1+c_2$};
\node (C) at (4,0.5) {$b_1+b_2$};
\node (C') at (4,-0.5) {$b_2+b_1$};
\node (D) at (2,-0.5) {$c_2+c_1$};
\node (E) at (5.5,0.5) {$d_{M_1 \otimes M_2} \in F(c_1+c_2)$};
\node (F) at (5.5,-0.5) {$d_{M_2 \otimes M_1} \in F(c_2+c_1)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1+i_2$} (B)
(C) edge node[above]{$o_1+o_2$} (B)
(A) edge node[left]{$\beta_{a_1,a_2}$} (A')
(C) edge node[left]{$\beta_{b_1,b_2}$} (C')
(A') edge node [above]{$i_2+i_1$} (D)
(C') edge node [above]{$o_2+o_1$} (D)
(B) edge node [left] {$\beta_{c_1,c_2}$} (D);
\end{tikzpicture}
\]
$$\iota_{\beta_{M_1,M_2}} \colon F(\beta_{c_1,c_2})(d_{M_1 \otimes M_2}) \xrightarrow{\sim} d_{M_2 \otimes M_1}$$ where the vertical 1-morphisms are given by braidings in $(\mathsf{A},+,0)$. This braiding makes the following triangle of underlying cospans commute:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$M_1 \otimes M_2$};
\node (B) at (2.25,1) {$M_1 \otimes M_2$};
\node (C) at (4.5,0) {$M_2 \otimes M_1$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$1$} (B)
(C) edge node[above]{$\beta_{M_2,M_1}$} (B)
(A) edge node[above]{$\beta_{M_1,M_2}$} (C);
\end{tikzpicture}
\]
as well as the following diagram of corresponding decorations in the category $F(c_1+c_2)$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$d_{M_1 \otimes M_2}$};
\node (B) at (2.25,1) {$d_{M_1 \otimes M_2} $};
\node (C) at (4.5,0) {$F(\beta_{c_2,c_1})(d_{M_2 \otimes M_1}) $};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$1$} (B)
(C) edge node[right]{$\iota_{\beta_{M_2,M_1}}$} (B)
(A) edge node[above]{$F(\beta_{c_2,c_1})(\iota_{\beta_{M_1,M_2}})$} (C);
\end{tikzpicture}
\]
since $F(\beta_{c_2,c_1} \beta_{c_1,c_2}) (d_{M_1 \otimes M_2}) = d_{M_1 \otimes M_2}$. Thus $F\mathbb{C}\textnormal{sp}_1$ is also symmetric monoidal.
